\begin{thebibliography}{99999}
\phantomsection
\singlespace\normalsize

\bibitem{CSA} Timothy I. Morgenthaler, MD, Vadim Kagramanov, MD; Viktor Hanak, MD; Paul A. Decker, MS, \textit{Complex Sleep Apnea Syndrome: Is It a Unique Clinical Syndrome?} SLEEP, Vol. 29, No. 9, 2006, \url{http://www.journalsleep.org/ViewAbstract.aspx?pid=26630}

\bibitem{NHLBI} \textit{What Are the Signs and Symptoms of Sleep Apnea?}, \url{http://www.nhlbi.nih.gov/health/health-topics/topics/sleepapnea/signs.html}

\bibitem{MMWR} Center for Disease Control and Prevention, \textit{Morbidity and Mortality Weekly Report}, March 2, 2011, \url{http://www.cdc.gov/mmwr/PDF/wk/mm6008.pdf}

\bibitem{accidents} Findley LJ, Unverzagt ME, Suratt PM., \textit{Automobile accidents involving patients with obstructive sleep apnea}, Department of Internal Medicine, University of Virginia, Charlottesville, August, 1988  \url{http://www.ncbi.nlm.nih.gov/pubmed/3195832}

\bibitem{accidents2} Barbe F. et. al \textit{Automobile Accidents in Patients with Sleep Apnea Syndrome}, American Journal of Respiratory and Critical Care Medicine, Vol. 158, No. 1 (1998), pp. 18-22, \url{http://www.atsjournals.org/doi/abs/10.1164/ajrccm.158.1.9709135}

\bibitem{statistics} Sleeping Disorder Statistics, \url{http://www.statisticbrain.com/sleeping-disorder-statistics/}

\bibitem{devices} Chunbai Zhang, MD, MPH, Mark Berger, MD, Atul Malhotra, MD, Stefanos N. Kales, MD, MPH, \textit{Portable Diagnostic Devices for Identifying Obstructive Sleep Apnea among Commercial Motor Vehicle Drivers: Considerations and Unanswered Questions}, SLEEP, Vol. 35, No. 11, 2012, \url{http://www.journalsleep.org/ViewAbstract.aspx?pid=28695}

\bibitem{PSG} Type I, Type II, Type III Sleep Monitors, CMS \& AASM Guidelines, and more \url{http://www.clevemed.com/terms/type-1_type-2_type-3_sleep_devices_by_AASM_and_CMS.shtml}

\bibitem{uvula} Lateral Pharyngoplasty \url{http://www.sleep-doctor.com/surgical-treatment-overview/palate-procedures/lateral-pharyngoplasty/}

\bibitem{surgery} Sleep surgery \url{http://en.wikipedia.org/wiki/Sleep_surgery}

\bibitem{oximeter} Pulse oximetry \url{http://en.wikipedia.org/wiki/Pulse_oximetry}

\bibitem{cpap} Sleep Apnea \url{http://www.macaulaydental.com/apnea.html}


\end{thebibliography}
